---
title: "COMP 112 Final Project"
output: html_document
date: "`r Sys.Date()`"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, eval = TRUE, warning = FALSE, message = FALSE, tidy = TRUE)
```

### Import Dataset & Activate Packages
```{r Loading packages and data}
library(broom)
library(dplyr)
library(ggplot2)
library(readr)
library(rpart.plot)
library(tidymodels) 
library(varhandle)
library(vip)
library(here)

tidymodels_prefer()
conflicted::conflict_prefer("vi", "vip")

set.seed(123)

tickets_base <- read_csv(here("res/tickets.csv"))
```

### Clean Data & Mutation
```{r Data cleaning}
tickets_mod <- tickets_base %>%
  select(-ticket, -ITOwner, -requestor) %>%
  rename(category = FiledAgainst) %>%
  rename(duration = daysOpen) %>%
  rename(impact = Severity) %>%
  rename(priority = Priority) %>%
  rename(satisfaction = Satisfaction) %>%
  rename(seniority = RequestorSeniority) %>%
  rename(type = TicketType) %>%
  mutate(category = replace(category, category == "Access/Login", "SSO/IAM")) %>%
  mutate(priority = replace(priority, priority == "0 - Unassigned", "Unclassified")) %>%
  mutate(priority = replace(priority, priority == "1 - Low", "Low")) %>%
  mutate(priority = replace(priority, priority == "2 - Medium", "Medium")) %>%
  mutate(priority = replace(priority, priority == "3 - High", "High")) %>%
  mutate(impact = replace(impact, impact == "0 - Unclassified", "Unclassified")) %>%
  mutate(impact = replace(impact, impact == "1 - Minor", "Single Person")) %>%
  mutate(impact = replace(impact, impact == "2 - Normal", "Group")) %>%
  mutate(impact = replace(impact, impact == "3 - Major", "Department")) %>%
  mutate(impact = replace(impact, impact == "4 - Critical", "Multiple Departments")) %>%
  mutate(satisfaction = replace(satisfaction, satisfaction == "0 - Unknown", "N/A")) %>%
  mutate(satisfaction = replace(satisfaction, satisfaction == "1 - Unsatisfied", "Unsatisfied")) %>%
  mutate(satisfaction = replace(satisfaction, satisfaction == "2 - Satisfied", "Satisfied")) %>%
  mutate(satisfaction = replace(satisfaction, satisfaction == "3 - Highly satisfied", "Highly Satisfied")) %>%
  mutate(seniority = replace(seniority, seniority == "1 - Junior", "Junior")) %>%
  mutate(seniority = replace(seniority, seniority == "2 - Regular", "Regular")) %>%
  mutate(seniority = replace(seniority, seniority == "3 - Senior", "Senior")) %>%
  mutate(seniority = replace(seniority, seniority == "4 - Management", "Management")) %>%
  mutate(type = replace(type, type == "Issue", "Malfunction")) 

# Implementation credit: Alex McCreight
tickets_mod2 <- tickets_mod %>% 
  mutate(wasEscalated = case_when(
    priority == "High" & impact == "Group" ~ 1,
    priority == "High" & impact == "Department" ~ 1,
    priority == "High" & impact == "Multiple Departments" ~ 1,
    TRUE ~ 0
  ))

str(tickets_mod2)

tickets <- as.data.frame(unclass(tickets_mod2), stringsAsFactors = TRUE)
tickets$wasEscalated <- as.factor(tickets$wasEscalated)
tickets$duration <- as.factor(tickets$duration)
```

## Research Question
What tickets are most likely to be escalated?

```{r}
tickets_escalated <- tickets_mod2 %>%
  filter(wasEscalated == 1)

# Lots of tickets from management and regular, less for senior, and the least for junior
# Interns probably won't be submitting a lot of tickets. 
# There are the most management and regular employees at any company
# Less senior staff in a company, so less tickets submitted.
ggplot(tickets_escalated) +
  geom_bar(aes(x = priority, fill = category)) +
  facet_wrap(vars(seniority))

# We can see that escalated tickets are priority High.
# Most escalated tickets are in the Systems category.
ggplot(tickets_mod2) +
  geom_bar(aes(x = priority, fill = category)) +
  facet_wrap(vars(wasEscalated))

# Tickets submitted by junior employees are rarely escalated.
# Management and regular have the highest count of escalated tickets, however senior tickets are also commonly escalated.
ggplot(tickets_mod2) +
  geom_bar(aes(x = priority, fill = seniority)) +
  facet_wrap(vars(wasEscalated))


```
Primary Question: What ticket characteristics are most helpful in predicting whether or not a ticket was escalated? How can we predict which tickets are/are not escalated? In other words, what tickets are most likely to be escalated?

- Seniority and Category are strong predictors as we have found out in our few initial visualizations. 

# Effective Visualization

```{r}
ggplot(tickets_mod2) +
  geom_bar(aes(x = priority, fill = seniority)) +
  facet_wrap(vars(wasEscalated))+theme_classic()

```

