---
title: "COMP 112 Final Project: Statistical Models"
author: "Ben Tang"
date: "`r Sys.Date()`"
html_document: default
pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, eval = TRUE, warning = FALSE, message = FALSE, tidy = TRUE)
```

## R Implementation
### Import Dataset & Activate Packages
```{r Loading packages and data}
library(broom)
library(dplyr)
library(ggplot2)
library(readr)
library(rpart.plot)
library(tidymodels) 
library(varhandle)
library(vip)
library(here)

tidymodels_prefer()
conflicted::conflict_prefer("vi", "vip")

set.seed(123)

tickets_base <- read_csv(here("res/tickets.csv"))
```

### Clean Data & Mutation
```{r Data cleaning}
tickets_mod <- tickets_base %>%
  select(-ticket, -ITOwner, -requestor) %>%
  rename(category = FiledAgainst) %>%
  rename(duration = daysOpen) %>%
  rename(impact = Severity) %>%
  rename(priority = Priority) %>%
  rename(satisfaction = Satisfaction) %>%
  rename(seniority = RequestorSeniority) %>%
  rename(type = TicketType) %>%
  mutate(category = replace(category, category == "Access/Login", "SSO/IAM")) %>%
  mutate(priority = replace(priority, priority == "0 - Unassigned", "Unclassified")) %>%
  mutate(priority = replace(priority, priority == "1 - Low", "Low")) %>%
  mutate(priority = replace(priority, priority == "2 - Medium", "Medium")) %>%
  mutate(priority = replace(priority, priority == "3 - High", "High")) %>%
  mutate(impact = replace(impact, impact == "0 - Unclassified", "Unclassified")) %>%
  mutate(impact = replace(impact, impact == "1 - Minor", "Single Person")) %>%
  mutate(impact = replace(impact, impact == "2 - Normal", "Group")) %>%
  mutate(impact = replace(impact, impact == "3 - Major", "Department")) %>%
  mutate(impact = replace(impact, impact == "4 - Critical", "Multiple Departments")) %>%
  mutate(satisfaction = replace(satisfaction, satisfaction == "0 - Unknown", "N/A")) %>%
  mutate(satisfaction = replace(satisfaction, satisfaction == "1 - Unsatisfied", "Unsatisfied")) %>%
  mutate(satisfaction = replace(satisfaction, satisfaction == "2 - Satisfied", "Satisfied")) %>%
  mutate(satisfaction = replace(satisfaction, satisfaction == "3 - Highly satisfied", "Highly Satisfied")) %>%
  mutate(seniority = replace(seniority, seniority == "1 - Junior", "Junior")) %>%
  mutate(seniority = replace(seniority, seniority == "2 - Regular", "Regular")) %>%
  mutate(seniority = replace(seniority, seniority == "3 - Senior", "Senior")) %>%
  mutate(seniority = replace(seniority, seniority == "4 - Management", "Management")) %>%
  mutate(type = replace(type, type == "Issue", "Malfunction")) 

# Implementation credit: Alex McCreight
tickets_mod2 <- tickets_mod %>% 
  mutate(wasEscalated = case_when(
    priority == "High" & impact == "Group" ~ 1,
    priority == "High" & impact == "Department" ~ 1,
    priority == "High" & impact == "Multiple Departments" ~ 1,
    TRUE ~ 0
  ))

str(tickets_mod2)

tickets <- as.data.frame(unclass(tickets_mod2), stringsAsFactors = TRUE)
tickets$wasEscalated <- as.factor(tickets$wasEscalated)
tickets$duration <- as.factor(tickets$duration)
```

## Regression Models
### Initial Fit
```{r Tickets model 1 initial fit}
tickets_mod1_lasso_spec <-
  linear_reg() %>%
  set_args(mixture = 1, penalty = 0) %>% 
  set_engine(engine = 'glmnet') %>%
  set_mode('regression')

tickets_mod1_lasso_recipe <- recipe(duration ~ ., data = tickets_mod2) %>%
    step_nzv(all_predictors()) %>%
    step_novel(all_nominal_predictors()) %>%
    step_normalize(all_numeric_predictors()) %>%
    step_dummy(all_nominal_predictors())

tickets_mod1_lasso_wf <- workflow() %>%
  add_recipe(tickets_mod1_lasso_recipe) %>%
  add_model(tickets_mod1_lasso_spec)

tickets_mod1_lasso_fit <- tickets_mod1_lasso_wf %>%
  fit(data = tickets_mod2)
```

### Cross-Validation and Final Fit
```{r Tickets model 1 cross-validation and final fit}
tickets_cv10_reg <- vfold_cv(tickets_mod2, v = 10)

tickets_mod1_lasso_spec_tune <-
  linear_reg() %>%
  set_args(mixture = 1, penalty = tune()) %>%
  set_engine(engine = 'glmnet') %>%
  set_mode('regression')

tickets_mod1_lasso_wf_tune <- workflow() %>%
  add_recipe(tickets_mod1_lasso_recipe) %>%
  add_model(tickets_mod1_lasso_spec_tune)

tickets_mod1_cv_penalty_grid <- grid_regular(
  penalty(range = c(-5, 3)),
  levels = 30)

tickets_mod1_tune_recipe <- tune_grid(
  tickets_mod1_lasso_wf_tune,
  resamples = tickets_cv10_reg,
  metrics = metric_set(mae),
  grid = tickets_mod1_cv_penalty_grid
)

tickets_mod1_tune_recipe %>% collect_metrics()

tickets_mod1_final_mod <- select_best(tickets_mod1_tune_recipe, metric = 'mae')

tickets_mod1_final_wf <- finalize_workflow(tickets_mod1_lasso_wf_tune, tickets_mod1_final_mod)

tickets_mod1_final_fit <- fit(tickets_mod1_final_wf, data = tickets_mod2)

tidy(tickets_mod1_final_fit)
```

### Residual Graphs
```{r Tickets residual graph}
tickets_mod1_final_output <- tickets_mod1_final_fit %>% 
    predict(new_data = tickets_mod2) %>%
    bind_cols(tickets_mod2) %>%
    mutate(resid = duration - .pred)

pdf("resid_graph.pdf",
    width = 5,
    height = 5)

ggplot(tickets_mod1_final_output, aes(x = .pred, y = resid)) +
    ggtitle("Residuals v. Predicted Values") +
    geom_point() +
    geom_smooth() +
    geom_hline(yintercept = 0, color = "red") +
    theme_classic() +
    labs(x = "Predicted Values", y = "Residuals")

dev.off()
```

#### Get Variable Importance
```{r}
glmnet_output <- tickets_mod1_final_fit %>% extract_fit_engine()
    
bool_predictor_exclude <- glmnet_output$beta==0

var_imp <- sapply(seq_len(nrow(bool_predictor_exclude)), function(row) {
    this_coeff_path <- bool_predictor_exclude[row,]
    if(sum(this_coeff_path) == ncol(bool_predictor_exclude)){ return(0)}else{
    return(ncol(bool_predictor_exclude) - which.min(this_coeff_path) + 1)}
})

var_imp_data <- tibble(
    var_name = rownames(bool_predictor_exclude),
    var_imp = var_imp
)
var_imp_data %>% arrange(desc(var_imp))
```

## Classification Models
### Model 1: Decision Tree
#### Pre-Processing
```{r CT Pre-processing}
tickets_cv10 <- vfold_cv(tickets, v = 10)

tickets_ct_tune <- decision_tree() %>%
  set_engine(engine = 'rpart') %>%
  set_args(cost_complexity = tune(),  
           min_n = 2, 
           tree_depth = NULL) %>% 
  set_mode('classification') 

tickets_rec <- recipe(wasEscalated ~ ., data = tickets) %>%
  step_rm(priority, impact) %>%
  step_dummy(all_nominal_predictors())

tickets_wf_tune <- workflow() %>%
  add_model(tickets_ct_tune) %>%
  add_recipe(tickets_rec)

cc_grid <- grid_regular(cost_complexity(range = c(-5, 1)), levels = 10)
```

#### Tuning Model
```{r CT Tune model, eval = FALSE}
tune_results <- tune_grid(
  tickets_wf_tune,
  resamples = tickets_cv10,
  grid = cc_grid,
  metrics = metric_set(accuracy)
)

save(tune_results, file = "tuneRes.RData")
```

#### Loading Tuned Model
```{r CT Load tuned model}
load("tuneRes.RData")
```

#### Visualize Best Cost Complexity
```{r CT Visualize best cost complexity}
pdf("cc_graph.pdf",
    width = 5,
    height = 5)

autoplot(tune_results) + 
  theme_classic() +
  geom_vline(xintercept = 0.1, color = "red", linetype = "dashed") +
  geom_vline(xintercept = 0.00001, color = "blue", linetype = "dashed") +
  geom_vline(xintercept = 0.00464, color = "purple", linetype = "dashed") +
  geom_vline(xintercept = 0.0002154435, color = "orange", linetype = "dashed") +
  labs(y = "Accuracy", title = "Cost Complexity vs. Accuracy")

dev.off()
```

#### Choose Final Model
```{r CT Choose final model}
best_cc <- select_by_one_std_err(tune_results, metric = "accuracy", desc(cost_complexity))

tickets_wf_final <- finalize_workflow(tickets_wf_tune, best_cc)
```

#### Fit Final Model
```{r CT Fit final CV model, eval = FALSE}
tickets_final_fit <- fit(tickets_wf_final, data = tickets)

save(tickets_final_fit, file = "cvFinalFit.RData")
```

#### Load Final Model and Collect Metrics
```{r CT Load final CV model}
load("cvFinalFit.RData")

tune_results %>%
  collect_metrics %>%
  filter(cost_complexity == best_cc %>% pull(cost_complexity))
```

#### Visualize Final Tree & Save to PDF
```{r Save final decision tree to PDF}
pdf("decision_tree.pdf",
    width = 5,
    height = 5)

tickets_ct <- tickets_final_fit %>% extract_fit_engine() %>% rpart.plot() 

dev.off()
```

#### Make Test Predictions
```{r CT Make predictions}
# Tickets 101 and 102 were escalated. Ticket 100 was not.
test_ticket_escN1 <- tickets[100, ]
test_ticket_escY1 <- tickets[101, ]
test_ticket_escY2 <- tickets[102, ]

# Soft
predict(tickets_final_fit, new_data = test_ticket_escN1, type = "prob")
predict(tickets_final_fit, new_data = test_ticket_escY1, type = "prob")
predict(tickets_final_fit, new_data = test_ticket_escY2, type = "prob")

# Hard
predict(tickets_final_fit, new_data = test_ticket_escN1, type = "class")
predict(tickets_final_fit, new_data = test_ticket_escY1, type = "class")
predict(tickets_final_fit, new_data = test_ticket_escY2, type = "class")
```

#### Evaluate CT Model
```{r CT Evaluate model}
metrics <- metric_set(sens, yardstick::spec, accuracy)

tickets_ct_output <- predict(tickets_final_fit, new_data = tickets) %>% bind_cols(tickets %>% select(wasEscalated))

tickets_ct_metrics <- tickets_ct_output %>%
  metrics(estimate = .pred_class, truth = wasEscalated)

tickets_final_fit %>%
  extract_fit_engine() %>%
  pluck('variable.importance')
```

### Model 2: Decision Tree & Bagging
#### Pre-Processing
```{r BM Pre-processing}
(tickets_rec %>% prep(tickets) %>% juice() %>% ncol()) - 1

tickets_bag_spec <- rand_forest() %>%
  set_engine(engine = 'ranger') %>%
  set_args(mtry = 58,
           trees = 1000,
           min_n = 2,
           probability = FALSE,
           importance = 'impurity') %>%
  set_mode('classification')

tickets_bag_wf <- workflow() %>%
  add_model(tickets_bag_spec) %>%
  add_recipe(tickets_rec)
```

#### Fit Model & Save Output
```{r BM Fit model, eval = FALSE}
tickets_bag_fit <- tickets_bag_wf %>%
  fit(data = tickets)

save(tickets_bag_fit, file = "ticketsBag_fit.RData")
```

#### Load Fitted Model & Get OOB Prediction Error
```{r BM Load fitted model}
load("ticketsBag_fit.RData")

# Get OOB prediction error: 31.48%
# tickets_bag_fit
```

#### Evaluate Model
```{r BM Evaluate model}
tickets_bag_OOB_out <- tibble(
  .pred_class = tickets_bag_fit %>% extract_fit_engine() %>% pluck('predictions'),
  wasEscalated = tickets %>% pull(wasEscalated))

tickets_bag_metrics <- metric_set(sens, yardstick::spec, accuracy)

tickets_bag_OOB_out %>%
  tickets_bag_metrics(truth = wasEscalated, estimate = .pred_class)

# Sens 77.93%
# Spec 51.49%
# Accu 68.49%
```

### Model 3: Bagging & Random Forest
#### Pre-Processing
```{r RF Pre-processing}
tickets_rf_spec <- rand_forest() %>%
  set_engine(engine = 'ranger') %>% 
  set_args(mtry = NULL,
           trees = 1000,
           min_n = 2,
           probability = FALSE,
           importance = 'impurity') %>% 
  set_mode('classification')

tickets_rf_wf <- workflow() %>%
  add_model(tickets_rf_spec) %>%
  add_recipe(tickets_rec)
```

#### Fit Model & Save Output
```{r RF Fit model, eval = FALSE}
tickets_rf_fit <- tickets_rf_wf %>%
  fit(data = tickets)

save(tickets_rf_fit, file = "ticketsRF_fit.RData")
```

#### Load Model & Get OOB Prediction Error
```{r RF Load model}
load("ticketsRF_fit.RData")

# Get OOB prediction error: 30.95%
# tickets_rf_fit
```

#### Evaluate Model Part 1: Metrics
```{r RF Evaluate model}
tickets_rf_OOB_out <- tibble(
  .pred_class = tickets_rf_fit %>% extract_fit_engine() %>% pluck('predictions'),
  wasEscalated = tickets %>% pull(wasEscalated), 
  seniority = tickets %>% pull(seniority))

tickets_rf_metrics <- metric_set(sens, yardstick::spec, accuracy)

tickets_rf_OOB_out %>% 
  tickets_rf_metrics(truth = wasEscalated, estimate = .pred_class)

# Sens: 78.02%
# Spec: 52.86%
# Accuracy: 69.04%
```

#### Evaluate Model Part 2: Confusion Matrix
```{r RF Evaluate model with confusion matrix}
tickets_rf_OOB_out %>% conf_mat(truth = wasEscalated, estimate = .pred_class)

tickets_rf_OOB_out %>% filter(wasEscalated == 1, .pred_class == 0) %>% count(seniority)

tickets %>% filter(seniority == "Regular") %>% count(seniority)
```

#### Evaluate Model Part 3a: (Save) Variable Importance Graphs
```{r RF Evaluate model with variable importance, eval = FALSE}
impurityVIP <- tickets_rf_fit %>% extract_fit_engine() %>% vip()

permutationVIP <- tickets_rf_wf %>%
  update_model(tickets_rf_spec %>% set_args(important = "permutation")) %>%
  fit(data = tickets) %>% extract_fit_engine() %>% vip()

save(impurityVIP, file = "impurityVIP.RData")
save(permutationVIP, file = "permutationVIP.RData")
```

#### Evaluate Model Part 3b: (Load) Variable Importance Graphs
```{r RF Visualize VIP graphs}
load("impurityVIP.RData")
load("permutationVIP.RData")

pdf("impurity_graph.pdf",
    width = 5,
    height = 5)

impurityVIP

pdf("permutation_graph.pdf",
    width = 5,
    height = 5)

permutationVIP

dev.off()
```